---
title: "ICC - Mixomics whole cohort"
author: "Miguel Cosenza & Tilman Werner - AG Schilling - Uniklinik Freiburg"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: null
  toc: yes
  toc_float: yes
  pdf_document: default
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width=5, fig.height=4, error=TRUE)
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
## Required packages ----
library(tidyverse)
library(readxl)
library(mixOmics)
library(kableExtra)
library(reactable)
library(proDA)
library(missForest)
library(dplyr)
library(tidyr)
library(ComplexHeatmap)
library(circlize)

## Required functions -----
na_filter <- function(data, max_na_per_group, logtrans, minInt_inpute, fraction_Inpute = 5){
      
data_long <- pivot_longer(data = data,
                             cols = matches("KB"),
                             values_to = "Intensity",
                             names_to = "Sample") 
data_long <- left_join(data_long, annotation1, by = "Sample")

if (logtrans){
       data_long <- mutate(data_long,
                           Intensity = log2(Intensity))
}

na_count <- group_by(data_long,
                     ID, Group) %>%
      summarise(na_count = sum(is.na(Intensity)),
                total = n()) %>% 
      ungroup() %>% 
      mutate(NA_fraction = na_count/total)

nafraction <- group_by(na_count,
                           ID) %>%
      summarise(max_nafraction = max(NA_fraction),
                min_nafraction = min(NA_fraction)) %>%
      ungroup()

included_prots <- dplyr::filter(nafraction,
                                max_nafraction <= max_na_per_group)



if (minInt_inpute){
  
  min_intensityprot <- data_long %>% 
  group_by(ID) %>% 
  mutate(min_int = min(Intensity, na.rm = TRUE)) %>%
  mutate(th_of_min_int = min_int/fraction_Inpute)
  
  data_long_nonas1 <- min_intensityprot %>% 
  mutate(Intensity = ifelse(is.na(Intensity),
                            yes = th_of_min_int,
                            no = Intensity),
         `Group-Sample` = paste(Group,"-",Sample, sep = "")) %>%
  dplyr::select(ID, `Group-Sample`, Intensity) %>%
  dplyr::filter(ID %in% included_prots$ID)

  data_maxperc_nas <- pivot_wider(data_long_nonas1,
                                  names_from = `Group-Sample`,
                                  values_from = Intensity)
} else {
  data_long_nonas1 <- filter(data_long,
                           ID %in% included_prots$ID) %>% 
      mutate(`Group-Sample` = paste(Group,"-",Sample, sep = "")) %>%
      dplyr::select(ID, `Group-Sample`, Intensity)
  
  data_maxperc_nas <- pivot_wider(data_long_nonas1,
                       names_from = `Group-Sample`,
                       values_from = Intensity)
}
}

## Function to get the proteins by their maximal contribution ----

getbycontribution <- function(data, splsda_model){
          
          n_comps <- splsda_model$ncomp
          
          prots_complist <- list()
          
          for (i in 1:n_comps){
                    prots_complist[[i]] <- selectVar(splsda_model, comp = i)$value %>% 
                              mutate(protein_id = row.names(.), comp = paste0("comp",i))
          }
          
          prots_percomp <- do.call(bind_rows, prots_complist)
          
          data_long <- pivot_longer(data = data,
                                    cols = matches("^S"),
                                    values_to = "Intensity",
                                    names_to = c("Group-Sample"))
          
          int_per_group <- separate(data_long, 
                                    col = `Group-Sample`, 
                                    sep = "-",
                                    into = c("Group", "Sample")) %>% 
                    group_by(ID, Group) %>% 
                    summarise(median_Int = median(Intensity),
                              mean_Int = mean(Intensity)) %>% 
                    filter(ID %in% prots_percomp$protein_id)
          
          proteins_group <- int_per_group %>%
                    ungroup() %>%
                    group_by(ID) %>%
                    filter(mean_Int == max(mean_Int)) %>% 
                    mutate(duplicated = duplicated(ID)) %>%
                    ungroup() %>% 
                    arrange(Group)
          
          return(proteins_group)
}

`%nin%` = Negate(`%in%`)
```

## Load data and general wrangling  

```{r}
## Load and prepare data and annotation ----
oridata <- read_excel("Data/Table S5 - ICC_proteomic_data_after_batch_correction.xlsx", na = "NA") %>%
  dplyr::rename(ID = `...1`)

annotation <- read_excel("Data/Table S1 - Patient Table.xlsx") %>%
              rename(Sample = 'Sample ID') %>%
              rename(Group = 'Tissue') %>%
              filter(Sample %nin% c("KB106", "KB42", "KB147", "KB53", "KB213", "KB215", "KB55", "KB23", "KB209")) %>%
              arrange(Sample)
annotation1 <- dplyr::select(annotation, Sample, Group)
predictors1 <- annotation1$Group

oridata1 <- dplyr::select(oridata, one_of("ID", annotation1$Sample)) %>%
  mutate_if(is.numeric, function(x, na.rm = TRUE) x - median(x, na.rm = TRUE))

# Load batch correction info
annot_batch <- read_excel("Data/annotation_batch_correction.xlsx") %>%
  rename(Sample = "FullRunName")

annotation <- left_join(annotation, annot_batch, by = "Sample")
```
Plot for ID numbers
```{r}
graph <- data.frame(colnames(oridata1[,-1]), colSums(!is.na(oridata1[,-1])), 1:ncol(oridata1[,-1])) 
colnames(graph) <- c("Sample", "IDnumber", "Samples") 

ggplot(graph, aes(Samples, sort(IDnumber))) +
  geom_point(size = 1) +
  labs(title = "Identified Protein Groups per Sample", y = "Protein Groups") +
  scale_y_continuous(limits = c(0, 3000))

ggplot(graph, aes(x = Samples, y = sort(IDnumber))) +
  geom_col(fill = "blanchedalmond", color = "black") +
  labs(y = "Identified Protein Groups") +
  scale_y_continuous(limits = c(0, 3000)) +
  theme_minimal() 
  # theme(axis.text=element_text(size=12),
  #       axis.title=element_text(size=14,face="bold"))

paste("average identified protein groups = ", mean(graph$IDnumber))
```

Plot for missingness across cohort
```{r}
expr_dat <- mutate(oridata1, sumNA = rowSums(is.na(oridata1)))  %>%
            mutate(completeness = (1 - (sumNA / (ncol(oridata1)-1))))

graph   <-  dplyr::select(expr_dat, c("ID", "sumNA", "completeness")) %>%
            arrange(-completeness) %>%
            mutate(ProteinGroups = 1:nrow(expr_dat))

lessthan80percent <- filter(graph, completeness >= 0.8)

graph$completeness <- (1-graph$completeness) * 100

ggplot(data = graph, aes(ProteinGroups, completeness)) +
  geom_line() +
  labs(title = "Missing Value Distribution", x = "Identified Protein Groups", y= "Missing Values [%]") +
  geom_hline(yintercept = 20, linetype = "dashed", color = "red") +
  theme_minimal() +
  scale_y_reverse()
```

# Data pre-processing  
For unsupervised statistics (clustering, PCA) an imputed data matrix is used, since these methods cannot process missing values. All other analyses rely on the original, unimputed matrix.
```{r}
data_nonaimpaft <- na_filter(data = oridata1,
                                max_na_per_group = 0.50,
                                logtrans = FALSE,
                                minInt_inpute = FALSE, 
                                fraction_Inpute = 1)

sample_nonaimpaft <- colnames(data_nonaimpaft)[-1]

### Prep expression matrix and transpose-----
tomat_noNAimpaft <-  data_nonaimpaft[,-1]

protein_idsimpaft <- data_nonaimpaft$ID

mat_noNAimpaft <- as.matrix(tomat_noNAimpaft)
row.names(mat_noNAimpaft) <- protein_idsimpaft
colnames(mat_noNAimpaft) <- sample_nonaimpaft

t_mat_nonaimpaft <- t(mat_noNAimpaft) 

### Log-2 transformation and median normalization  ----

#data_1st_log2_med_noraft <- apply(X = mat_noNAimpaft, MARGIN = 2, FUN = log2)
data_1st_log2_med_noraft <- median_normalization(mat_noNAimpaft)
t_data_1st_log2_med_normaft <- t(data_1st_log2_med_noraft)

data_1st_log2_med_noraftid <- rownames_to_column(as.data.frame(data_1st_log2_med_noraft), "Protein")
data_1st_log2_med_noraftid1 <- rownames_to_column(as.data.frame(data_1st_log2_med_noraft), "ID")


### Random forest missing value imputation ----

if(!file.exists("results/imputfr_aft.rds")){
  set.seed(546)
  #tictoc::tic()
  impfr_2 <- missForest(t_data_1st_log2_med_normaft) # this takes ~ 3 minutes to compute
  #tictoc::toc()
  write_rds(impfr_2, file = "results/imputfr_aft.rds")
} else {
  impfr_2 <- read_rds(file = "results/imputfr_aft.rds")
}

t_data_1st_log2_med_normimpaft <- impfr_2$ximp

data_1st_log2_med_normimpaft <- t(t_data_1st_log2_med_normimpaft)
```   

# Principal component analysis (PCA)  
```{r}
pca_res = pca(t_data_1st_log2_med_normimpaft, ncomp = 10, center = TRUE, scale = TRUE)

plot(pca_res)

plotIndiv(pca_res, group = predictors1, ind.names = FALSE, 
          legend = TRUE, title = 'PCA results', col.per.group = c("cadetblue", "red"))
```   

# Hierarchical Clustering   
```{r, fig.width=12, fig.height=6, results="asis"}
my_dist_t <- dist(t_data_1st_log2_med_normimpaft)
my_hclust_t <- hclust(my_dist_t)
plot(my_hclust_t)
```

#Heatmap
Indicating measurement and digestion batches
```{r}
oridata2 <- column_to_rownames(oridata1, var = "ID")
oridata2 <- filter(oridata2, rowSums(!is.na(oridata2)) / ncol(oridata2) > 0.8) %>%
  dplyr::select(annotation$Sample) 
subtractor <- apply(oridata2[,-1], 1, median, na.rm = TRUE)
oridata2 <- oridata2 - subtractor
oridata2 <- as.matrix(oridata2) 

annotation$Group <- str_replace_all(annotation$Group, "normal", "TANM")

#CAVE! annotation file and plot matrix need to be in the exactly same order to match in the plot!
ha = columnAnnotation(Tissue = annotation$Group,
                      digest_batch = annotation$digestion_batch,
                      MS_batch = annotation$MS_batch,
                      measurement_order = anno_lines(annotation$measurement_order),
                      col = list(Tissue = c("Tumor" = "red", "TANM" = "lightblue"),
                                 MS_batch = colorRamp2(c(1, 11, 21), c("green", "yellow", "red")),
                                 digest_batch = colorRamp2(c(1, 5, 10), c("green", "yellow", "red"))))

heatmap <- Heatmap(oridata2, name = "∆ log2 LFQ", column_title = "Samples", row_title = "Proteins", 
        show_row_names = FALSE, show_column_names = FALSE, column_names_gp = gpar(fontsize = 5),
        show_row_dend = FALSE,
        cluster_columns = my_hclust_t,
        bottom_annotation = ha,
        heatmap_legend_param = list(title = "∆ log2 LFQ", title_position = "lefttop-rot"))

heatmap

tiff("Heatmap.tiff", units = "in", width = 6, height = 8, res = 200, pointsize =(8*100/72))
heatmap
dev.off()
```

# Immunemap
Including custom-made list of Immune Cell Markers
```{r}
oridata_immune <- column_to_rownames(oridata, var = "ID") %>%
  filter(rowSums(!is.na(.)) / ncol(.) > 0.3) %>%
  dplyr::select(annotation1$Sample)
subtractor <- apply(oridata_immune, 1, median, na.rm = TRUE)
oridata_immune <- oridata_immune - subtractor
oridata_immune <- rownames_to_column(oridata_immune, var = "UniProt_ID_Human")

immunelist <- read_excel("Data/CDMarkers_Profiling_ImmuneCells_Human_Mouse.xlsx") 
immunelist1 <- filter(immunelist, !is.na(`Key Marker`))

oridata_immune1 <- filter(oridata_immune, 
                          UniProt_ID_Human %in% immunelist1$UniProt_ID_Human) %>%
  left_join(dplyr::select(immunelist1, "UniProt_ID_Human", "Gene"), by = "UniProt_ID_Human") %>%
  column_to_rownames(var = "Gene") %>%
  dplyr::select(-UniProt_ID_Human) %>%
  as.matrix()

annotation_immune <- mutate(annotation, Group = factor(Group,
                                                            levels = c("Tumor", "TANM")))

immunelist_fit <- filter(immunelist1, Gene %in% rownames(oridata_immune1))
```

```{r}
immunelist2 <- filter(immunelist, !is.na(Gene))

oridata_immune2 <- filter(oridata_immune, 
                          UniProt_ID_Human %in% immunelist2$UniProt_ID_Human) %>%
  inner_join(dplyr::select(immunelist2, "UniProt_ID_Human", "Gene"), by = "UniProt_ID_Human") %>%
  unique() %>%
  remove_rownames() %>%
  column_to_rownames(var = "Gene") %>%
  dplyr::select(-UniProt_ID_Human) %>%
  as.matrix()

annotation_immune2 <- mutate(annotation, Group = factor(Group,
                                                            levels = c("Tumor", "TANM")))

immunelist_fit2 <- filter(immunelist2, Gene %in% rownames(oridata_immune2))
```

```{r}
#CAVE! annotation file and plot matrix need to be in the exactly same order to match in the plot!
ha = columnAnnotation(Tissue = annotation_immune$Group,
                   col = list(Tissue = c("Tumor" = "red", "TANM" = "lightblue")),
                   show_legend = FALSE)

hb = rowAnnotation(Category = immunelist_fit$`Key Marker`,
                   show_annotation_name = FALSE, 
                   show_legend = FALSE)

Heatmap <- Heatmap(oridata_immune2, name = "∆ log2 LFQ", column_title = " ", row_title = "", 
        col = colorRamp2(c(-5, 0, 5), c("blue", "grey93", "red")),
        show_row_names = TRUE, show_column_names = FALSE, 
        row_names_side = "left", row_names_gp = gpar(fontsize = 5),
        show_row_dend = FALSE, show_column_dend = FALSE, 
        cluster_rows = TRUE, cluster_columns = FALSE,
        column_split = annotation_immune$Group, 
        #row_split = immunelist_fit$`Key Marker`, 
        row_title_rot = 0, 
        row_title_gp = gpar(fontsize = 5), row_title_side = "right",
        bottom_annotation = ha,  
        #right_annotation = hb,
        na_col = "white",
        heatmap_legend_param = list(title = "∆ log2 LFQ", title_position = "lefttop-rot"))
Heatmap

tiff("Heatmap_immune2.tiff", units = "in", width = 5, height = 5, res = 300, pointsize =(5*100/72))
Heatmap
dev.off()
```