---
title: "ICC - Mixomics"
author: "Miguel Cosenza, Tilman Werner - AG Schilling - Uniklinik Freiburg"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width=5, fig.height=4, error=TRUE)
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
## Required packages ----
library(tidyverse)
library(readxl)
library(mixOmics)
library(kableExtra)
library(reactable)
library(proDA)
library(missForest)
library(dplyr)
library(M3C)
library(made4)
library(ComplexHeatmap)
library(circlize)
library(lubridate)
library(patchwork)
library(rcompanion)
`%notin%` <- Negate(`%in%`)

## Required functions -----
na_filter <- function(data, max_na_per_group, logtrans, minInt_inpute, fraction_Inpute = 5){
      
data_long <- pivot_longer(data = data,
                             cols = matches("KB"),
                             values_to = "Intensity",
                             names_to = "Sample") 
data_long <- left_join(data_long, annotation1, by = "Sample")

if (logtrans){
       data_long <- mutate(data_long,
                           Intensity = log2(Intensity))
}

na_count <- group_by(data_long,
                     ID, Group) %>%
      summarise(na_count = sum(is.na(Intensity)),
                total = n()) %>% 
      ungroup() %>% 
      mutate(NA_fraction = na_count/total)

nafraction <- group_by(na_count,
                           ID) %>%
      summarise(max_nafraction = max(NA_fraction),
                min_nafraction = min(NA_fraction)) %>%
      ungroup()

included_prots <- dplyr::filter(nafraction,
                                max_nafraction <= max_na_per_group)



if (minInt_inpute){
  
  min_intensityprot <- data_long %>% 
  group_by(ID) %>% 
  mutate(min_int = min(Intensity, na.rm = TRUE)) %>%
  mutate(th_of_min_int = min_int/fraction_Inpute)
  
  data_long_nonas1 <- min_intensityprot %>% 
  mutate(Intensity = ifelse(is.na(Intensity),
                            yes = th_of_min_int,
                            no = Intensity),
         `Group-Sample` = paste(Group,"-",Sample, sep = "")) %>%
  dplyr::select(ID, `Group-Sample`, Intensity) %>%
  dplyr::filter(ID %in% included_prots$ID)

  data_maxperc_nas <- pivot_wider(data_long_nonas1,
                                  names_from = `Group-Sample`,
                                  values_from = Intensity)
} else {
  data_long_nonas1 <- filter(data_long,
                           ID %in% included_prots$ID) %>% 
      mutate(`Group-Sample` = paste(Group,"-",Sample, sep = "")) %>%
      dplyr::select(ID, `Group-Sample`, Intensity)
  
  data_maxperc_nas <- pivot_wider(data_long_nonas1,
                       names_from = `Group-Sample`,
                       values_from = Intensity)
}
}

## Function to get the proteins by their maximal contribution ----

getbycontribution <- function(data, splsda_model){
          
          n_comps <- splsda_model$ncomp
          
          prots_complist <- list()
          
          for (i in 1:n_comps){
                    prots_complist[[i]] <- selectVar(splsda_model, comp = i)$value %>% 
                              mutate(protein_id = row.names(.), comp = paste0("comp",i))
          }
          
          prots_percomp <- do.call(bind_rows, prots_complist)
          
          data_long <- pivot_longer(data = data,
                                    cols = matches("^S"),
                                    values_to = "Intensity",
                                    names_to = c("Group-Sample"))
          
          int_per_group <- separate(data_long, 
                                    col = `Group-Sample`, 
                                    sep = "-",
                                    into = c("Group", "Sample")) %>% 
                    group_by(ID, Group) %>% 
                    summarise(median_Int = median(Intensity),
                              mean_Int = mean(Intensity)) %>% 
                    filter(ID %in% prots_percomp$protein_id)
          
          proteins_group <- int_per_group %>%
                    ungroup() %>%
                    group_by(ID) %>%
                    filter(mean_Int == max(mean_Int)) %>% 
                    mutate(duplicated = duplicated(ID)) %>%
                    ungroup() %>% 
                    arrange(Group)
          
          return(proteins_group)
}

`%nin%` = Negate(`%in%`)
```

## Load and prepare data and annotation 
```{r}
oridata <- read_excel("Data/Table S5 - ICC_proteomic_data_after_batch_correction.xlsx", na = "NA") %>%
  dplyr::rename(ID = `...1`)

annotation <- read_excel("Data/Table S1 - Patient Table.xlsx") %>%
              dplyr::rename(Sample = 'Sample ID') %>%
              dplyr::rename(Group = 'Tissue') %>%
              filter(Sample %nin% c("KB106", "KB42", "KB147", "KB53", "KB213", "KB215", "KB55", "KB23", "KB209")) %>%
              arrange(Sample)
annotation1 <- dplyr::select(annotation, Sample, Group)
annotation2 <- filter(annotation, Group == "Tumor") 

oridata1 <- dplyr::select(oridata, "ID", annotation2$Sample) %>%
  mutate_if(is.numeric, function(x, na.rm = TRUE) x - median(x, na.rm = TRUE))
```

# Data pre-processing  
For unsupervised statistics (clustering, PCA) an imputed data matrix is used, since these methods cannot process missing values. All other analyses rely on the original, unimputed matrix.
```{r}
data_nonaimpaft <- na_filter(data = oridata1,
                                max_na_per_group = 0.50,
                                logtrans = FALSE,
                                minInt_inpute = FALSE, 
                                fraction_Inpute = 1)

sample_nonaimpaft <- colnames(data_nonaimpaft)[-1]

### Prep expression matrix and transpose-----
tomat_noNAimpaft <-  data_nonaimpaft[,-1]

protein_idsimpaft <- data_nonaimpaft$ID

mat_noNAimpaft <- as.matrix(tomat_noNAimpaft)
row.names(mat_noNAimpaft) <- protein_idsimpaft
colnames(mat_noNAimpaft) <- sample_nonaimpaft

t_mat_nonaimpaft <- t(mat_noNAimpaft) 

data_1st_log2_med_noraft <- apply(X = mat_noNAimpaft, MARGIN = 2, FUN = log2)
data_1st_log2_med_noraft <- median_normalization(mat_noNAimpaft)
t_data_1st_log2_med_normaft <- t(data_1st_log2_med_noraft)

data_1st_log2_med_noraftid <- rownames_to_column(as.data.frame(data_1st_log2_med_noraft), "Protein")
data_1st_log2_med_noraftid1 <- rownames_to_column(as.data.frame(data_1st_log2_med_noraft), "ID")


### Random forest missing value imputation ----

if(!file.exists("results/imputfr_aft_intratumor.rds")){
  set.seed(546)
  #tictoc::tic()
  impfr_2 <- missForest(t_data_1st_log2_med_normaft) # this takes ~ 3 minutes to compute
  #tictoc::toc()
  write_rds(impfr_2, file = "results/imputfr_aft_intratumor.rds")
} else {
  impfr_2 <- read_rds(file = "results/imputfr_aft_intratumor.rds")
}

t_data_1st_log2_med_normimpaft <- impfr_2$ximp

data_1st_log2_med_normimpaft <- t(t_data_1st_log2_med_normimpaft)

```   

# Hierarchical Clustering   
```{r, fig.width=12, fig.height=6, results="asis"}
my_dist_t <- dist(t_data_1st_log2_med_normimpaft)
my_hclust_t <- hclust(my_dist_t)
plot(my_hclust_t)
```

# M3C determination of most appropiate number of subgroups in cluster
```{r fig.width=12, fig.height=6, results="asis", echo=FALSE, include=FALSE}
m3c <- M3C(data_1st_log2_med_normimpaft)
m3c
```


# Principal component analysis (PCA)
```{r}
pca_res = mixOmics::pca(t_data_1st_log2_med_normimpaft, ncomp = 10, center = TRUE, scale = TRUE)

plot(pca_res)
plotIndiv(pca_res, ind.names = FALSE)

pca <- plotIndiv(pca_res, ind.names = FALSE)
pca_df <- pca$df %>%
  rownames_to_column(var = "Sample") %>%
  mutate(Sample = str_remove_all(Sample, "Tumor-")) %>%
  inner_join(annotation, by = "Sample") 

plot1 <- ggplot(pca_df, aes(x, y, color = cluster)) +
  geom_point(size = 3) +
  theme_classic() +
  xlab(pca$graph$labels$x) +
  ylab(pca$graph$labels$y) +
  labs(title = "PCA clusters") +
  scale_color_manual(values = c("blue", "orange"),
                     breaks = c("cluster1", "cluster2"))
plot1
```   

#Heatmap
```{r}
annot_dat <- annotation2 %>%
  dplyr::rename(ID = 'Sample') %>%
  filter(ID %notin% c("KB106", "KB42", "KB147", "KB53", "KB213", "KB215", "KB55", "KB23", "KB209")) %>%
  arrange(ID) %>%
  mutate(daystorecurrence = as.numeric(daystorecurrence))
annot_dat$`Tumor Stage` <- factor(annot_dat$`Tumor Stage`, levels = c("T1", "T2", "T3", "T4"))


oridata2 <- column_to_rownames(oridata1, var = "ID")
oridata2 <- filter(oridata2, rowSums(!is.na(oridata2)) / ncol(oridata2) > 0.8)
subtractor <- apply(oridata2[,-1], 1, median, na.rm = TRUE)
oridata2 <- oridata2 - subtractor
oridata2 <- as.matrix(oridata2) 

#CAVE! annotation file and plot matrix need to be in the exactly same order to match in the plot!
ha = columnAnnotation(cluster = annot_dat$cluster, 
                      TTR = annot_dat$daystorecurrence, 
                      Sex = annot_dat$Sex, 
                      Radiotherapy = annot_dat$Radiotherapy,
                      DuctType = annot_dat$`Duct Type`,
                      TumorStage = annot_dat$`Tumor Stage`,
                   col = list(cluster = c("cluster1" = "cornflowerblue", "cluster2" = "goldenrod2"), 
                              TTR = colorRamp2(c(0, 1300, 2300), c("red", "yellow", "green")),
                              Sex = c("M" = "blue", "F" = "pink"),
                              Radiotherapy = c("Yes" = "red", "No" = "gray95"),
                              DuctType = c("LDT" = "#D29BA2", "SDT" = "#8E698E"),
                              TumorStage = c("T1" = "#FFF0F0", "T2" = "#FF9595", "T3" = "#DD3730", "T4" = "#9b0303")), 
                   na_col = "white",
                   annotation_label = c("Tumor Cluster", "TTR (days)", "Sex", "Radiotherapy", 
                                        "Duct Type", "AJCC Tumor Stage"),
                   annotation_legend_param = list(cluster = list(title_position = "lefttop-rot"),
                                                  Radiotherapy = list(title_position = "lefttop-rot"),
                                                  Sex = list(title_position = "lefttop-rot"),
                                                  DuctType = list(title_position = "lefttop-rot"),
                                                  TumorStage = list(title_position = "lefttop-rot"),
                                                  TTR = list(title_position = "lefttop-rot")))

Heatmap(oridata2, name = "∆ log2 LFQ", column_title = "Samples", row_title = "Proteins", 
        show_row_names = FALSE, show_column_names = TRUE, 
        show_row_dend = FALSE,
        cluster_columns = my_hclust_t,
        bottom_annotation = ha,
        heatmap_legend_param = list(title = "∆ log2 LFQ", title_position = "lefttop-rot"))
```

# Correlation Plot
```{r}
corr <- column_to_rownames(oridata1, var = "ID") %>%
  cor(use = "pairwise.complete.obs", method = "pearson")

#CAVE! annotation file and plot matrix need to be in the exactly same order to match in the plot!
ha = columnAnnotation(cluster = annot_dat$cluster, 
                      TTR = annot_dat$daystorecurrence, 
                      Sex = annot_dat$Sex, 
                      Radiotherapy = annot_dat$Radiotherapy,
                      DuctType = annot_dat$`Duct Type`,
                      TumorStage = annot_dat$`Tumor Stage`,
                   col = list(cluster = c("cluster1" = "cornflowerblue", "cluster2" = "goldenrod2"), 
                              TTR = colorRamp2(c(0, 1300, 2300), c("red", "yellow", "green")),
                              Sex = c("M" = "blue", "F" = "pink"),
                              Radiotherapy = c("Yes" = "red", "No" = "gray95"),
                              DuctType = c("LDT" = "#D29BA2", "SDT" = "#8E698E"),
                              TumorStage = c("T1" = "#FFF0F0", "T2" = "#FF9595", "T3" = "#DD3730", "T4" = "#9b0303")), 
                   na_col = "white",
                   annotation_label = c("Tumor Cluster", "TTR (days)", "Sex", "Radiotherapy", 
                                        "Duct Type", "AJCC Tumor Stage"),
                   annotation_legend_param = list(cluster = list(title_position = "lefttop-rot"),
                                                  Radiotherapy = list(title_position = "lefttop-rot"),
                                                  Sex = list(title_position = "lefttop-rot"),
                                                  DuctType = list(title_position = "lefttop-rot"),
                                                  TumorStage = list(title_position = "lefttop-rot"),
                                                  TTR = list(title_position = "lefttop-rot")))

Heatmap(corr,  
        col = colorRamp2(c(0.7, 0.85, 0.95,  1), c("grey", "white", "red", "red4")), 
        #col = colorRamp2(c(0.75,  1), c("white", "black")),
        show_row_names = FALSE, show_column_names = FALSE, 
        show_row_dend = FALSE,
        cluster_columns = my_hclust_t, cluster_rows = my_hclust_t,
        column_names_gp = gpar(fontsize = 5), row_names_gp = gpar(fontsize = 5),
        bottom_annotation = ha,
        heatmap_legend_param = list(title = "correlation", title_position = "lefttop-rot"))
```

# Immunemap
```{r}
oridata_immune <- column_to_rownames(oridata1, var = "ID") %>%
  filter(rowSums(!is.na(.)) / ncol(.) > 0.3) 
subtractor <- apply(oridata_immune, 1, median, na.rm = TRUE)
oridata_immune <- oridata_immune - subtractor
oridata_immune <- rownames_to_column(oridata_immune, var = "UniProt_ID_Human")

immunelist <- read_excel("Data/CDMarkers_Profiling_ImmuneCells_Human_Mouse.xlsx") 
immunelist1 <- filter(immunelist, !is.na(`Key Marker`))

oridata_immune1 <- filter(oridata_immune, 
                          UniProt_ID_Human %in% immunelist1$UniProt_ID_Human) %>%
  left_join(dplyr::select(immunelist1, "UniProt_ID_Human", "Gene"), by = "UniProt_ID_Human") %>%
  column_to_rownames(var = "Gene") %>%
  dplyr::select(-UniProt_ID_Human) %>%
  as.matrix()

annotation_immune <- mutate(annotation2, cluster = factor(cluster,
                                                            levels = c("cluster1", "cluster2")))

immunelist_fit <- filter(immunelist1, Gene %in% rownames(oridata_immune1))
```

```{r}
#CAVE! annotation file and plot matrix need to be in the exactly same order to match in the plot!
ha = columnAnnotation(cluster = annotation_immune$cluster,
                   col = list(cluster = c("cluster1" = "cornflowerblue", "cluster2" = "goldenrod2")), 
                   show_legend = FALSE)

hb = rowAnnotation(Category = immunelist_fit$`Key Marker`,
                   show_annotation_name = FALSE, 
                   show_legend = FALSE)

Heatmap <- Heatmap(oridata_immune1, name = "∆ log2 LFQ", column_title = " ", row_title = "", 
        col = colorRamp2(c(-5, 0, 5), c("blue", "grey93", "red")),
        show_row_names = TRUE, show_column_names = FALSE, 
        row_names_side = "left", row_names_gp = gpar(fontsize = 5),
        show_row_dend = FALSE, show_column_dend = FALSE, 
        cluster_rows = FALSE, cluster_columns = FALSE,
        column_split = annotation_immune$cluster, 
        row_split = immunelist_fit$`Key Marker`, row_title_rot = 0, 
        row_title_gp = gpar(fontsize = 5), row_title_side = "right",
        bottom_annotation = ha, right_annotation = hb, na_col = "white",
        heatmap_legend_param = list(title = "∆ log2 LFQ", title_position = "lefttop-rot"))
Heatmap

tiff("Heatmap_immune.tiff", units = "in", width = 6, height = 3, res = 200, pointsize =(6*100/72))
Heatmap
dev.off()
```


```{r}
immunelist2 <- filter(immunelist, !is.na(Gene))

oridata_immune2 <- filter(oridata_immune, 
                          UniProt_ID_Human %in% immunelist2$UniProt_ID_Human) %>%
  inner_join(dplyr::select(immunelist2, "UniProt_ID_Human", "Gene"), by = "UniProt_ID_Human") %>%
  unique() %>%
  remove_rownames() %>%
  column_to_rownames(var = "Gene") %>%
  dplyr::select(-UniProt_ID_Human) %>%
  as.matrix()

annotation_immune2 <- mutate(annotation2, cluster = factor(cluster,
                                                            levels = c("cluster1", "cluster2")))

immunelist_fit2 <- filter(immunelist2, Gene %in% rownames(oridata_immune2))
```

```{r}
#CAVE! annotation file and plot matrix need to be in the exactly same order to match in the plot!
ha = columnAnnotation(cluster = annotation_immune$cluster,
                   col = list(cluster = c("cluster1" = "cornflowerblue", "cluster2" = "goldenrod2")), 
                   show_legend = FALSE)

hb = rowAnnotation(Category = immunelist_fit$`Key Marker`,
                   show_annotation_name = FALSE, 
                   show_legend = FALSE)

Heatmap <- Heatmap(oridata_immune2, name = "∆ log2 LFQ", column_title = " ", row_title = "", 
        col = colorRamp2(c(-5, 0, 5), c("blue", "grey93", "red")),
        show_row_names = TRUE, show_column_names = FALSE, 
        row_names_side = "left", row_names_gp = gpar(fontsize = 5),
        show_row_dend = FALSE, show_column_dend = FALSE, 
        cluster_rows = TRUE, cluster_columns = FALSE,
        column_split = annotation_immune$cluster, 
        #row_split = immunelist_fit$`Key Marker`, 
        row_title_rot = 0, 
        row_title_gp = gpar(fontsize = 5), row_title_side = "right",
        bottom_annotation = ha,  
        #right_annotation = hb,
        na_col = "white",
        heatmap_legend_param = list(title = "∆ log2 LFQ", title_position = "lefttop-rot"))
Heatmap

tiff("Heatmap_immune2.tiff", units = "in", width = 10, height = 10, res = 200, pointsize =(10*100/72))
Heatmap
dev.off()
```

# Matrisomal heatmap
```{r}
matrixprot <- read_excel("Data/Kopie von matrisome_hs_masterlist-1.xlsx") 
matrixprot2 <- matrixprot 
matrixprot2$UniProt_IDs <- str_split_fixed(matrixprot$UniProt_IDs, ":", n = Inf) 

oridat_matrisom <- filter(oridata1, ID %in% matrixprot2$UniProt_IDs) %>%
  dplyr::select("ID", annot_dat$ID)

oridat_matrisom2 <- column_to_rownames(oridat_matrisom, var = "ID")
oridat_matrisom2 <- filter(oridat_matrisom2, rowSums(!is.na(oridat_matrisom2)) / ncol(oridat_matrisom2) > 0.8)
subtractor <- apply(oridat_matrisom2[,-1], 1, median, na.rm = TRUE)
oridat_matrisom2 <- oridat_matrisom2 - subtractor


matrixprot_res <- sapply(matrixprot$UniProt_IDs, str_extract, rownames(oridat_matrisom2)) %>%
  t()%>%
  as.data.frame()

i <- 1
blabla <- vector()
while (i <= nrow(matrixprot_res)) {
  x <- t(matrixprot_res[i,])
  bla <- ifelse(sum(!is.na(x)) > 0, na.omit(x), NA)
  blabla <- c(blabla, bla)
  i <- i + 1
}

matrixprot3 <- mutate(matrixprot2, detect = blabla) %>%
  filter(!is.na(detect))

oridat_matrisom3 <- rownames_to_column(oridat_matrisom2, var = "detect") %>%
  left_join(matrixprot3, by = "detect")

oridat_matrisom4 <- dplyr::select(oridat_matrisom3, "detect", contains("KB")) %>%
  column_to_rownames(var = "detect")

matrixprot4 <- dplyr::select(oridat_matrisom3, !contains("KB")) 


ha = columnAnnotation(cluster = annot_dat$cluster,
                      col = list(cluster = c("cluster1" = "cornflowerblue", "cluster2" = "goldenrod2")),
                      show_annotation_name = FALSE,
                      annotation_legend_param = list(title = "cluster", title_position = "lefttop-rot"))

hb = rowAnnotation(Category = matrixprot4$Category, 
                   col = list(Category = c("Collagens" = "#235781", "ECM-affiliated Proteins" = "#A25525",
                                           "ECM Glycoproteins" = "#A8D8CA", "ECM Regulators" = "#E8A470",
                                           "Proteoglycans" = "#711550", "Secreted Factors" = "#F0E7CD")),
                   show_annotation_name = FALSE, 
                   annotation_legend_param = list(title = "Category", title_position = "lefttop-rot"))

ht <- Heatmap(oridat_matrisom4, name = "∆ log2 LFQ",
        column_title = "Samples", row_title = "  ", 
        show_row_names = TRUE, show_column_names = FALSE, 
        na_col = "white",
        column_names_gp = gpar(fontsize = 5), row_names_gp = gpar(fontsize = 6),
        show_row_dend = FALSE,
        row_split = matrixprot4$Category,
        cluster_rows = TRUE,
        row_labels = matrixprot4$`Gene Symbol`,
        cluster_columns = my_hclust_t,
        top_annotation = ha,
        right_annotation = hb,
        heatmap_legend_param = list(title = "∆ log2 LFQ", title_position = "lefttop-rot"))
ht
```
