---
title: "ICC_Cox_regression"
author: "Tilman Werner"
date: "5/14/2021"
output: html_document
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(survival)
library(survminer)
library(dplyr)
library(grDevices)
library(knitr)
library(rstatix)
library(ggpubr)
library(DOSE)
library(org.Hs.eg.db)
library(enrichplot)
library(ggnewscale)
library(ggupset)
library(clusterProfiler)
library(made4)
library(ggridges)
library(readxl)
library(gtsummary)
`%notin%` <- Negate(`%in%`)
```


```{r, echo=FALSE, include=FALSE}
### Load and modify data
wide_dat_pre <- read_excel("Data/Table S5 - ICC_proteomic_data_after_batch_correction.xlsx", na = "NA") %>%
  dplyr::rename(ID = `...1`)
```

#### Load annotation file
```{r}
# prepare annotation for entire cohort 
annot_dat <- read_excel("Data/Table S1 - Patient Table.xlsx") %>%
              dplyr::rename(Sample = 'Sample ID') %>%
              filter(Sample %notin% c("KB106", "KB42", "KB147", "KB53", "KB213", "KB215", "KB55", "KB23", "KB209")) %>%
              filter(Tissue == "Tumor") %>%
              arrange(Sample) %>%
              mutate(`Tumor Stage` = factor(`Tumor Stage`, levels = c("T1", "T2", "T3", "T4"))) %>%
              mutate(Radiotherapy = factor(Radiotherapy, levels = c("Yes", "No"))) %>%
              mutate(cluster = factor(cluster, levels = c("cluster1", "cluster2")))
```


### Define timepoints
The clinical annotation includes three timepoints to consider: the tumor resection date, the date of the first recurrence and the last follow-up date.  Days-to-recurrence are the time difference between resection and first recurrence date - days-to-last-follow-up were calculated in the same fashion. 
For log-rank statistics, patients who did not show a recurrence throughout the course of the study have to be included and censored. Censoring means that they will be considered for statistics, although they obviously did not show a recurrence. Since we know that these patients were recurrence-free until their last follow-up date, we can only include them in the study for the time between their tumor resection and the last follow-up. After their last recurrence free follow up these patients are removed from the log-rank statistics (a.k.a. censoring) without impacting the survival rate. Furthermore, the study is defined to end with the longest time to recurrence. All patients whose time to last follow-up is longer will be censored from then on.  

```{r, echo=FALSE, include=FALSE}
# Filter samples according to annotation & median normalizing
wide_dat <- dplyr::select(wide_dat_pre, ID, one_of(annot_dat$Sample))
wide_dat <- mutate_if(wide_dat, is.numeric, function(x, na.rm = TRUE) x - median(x, na.rm = TRUE))
```

# Distribution of time to recurrence across samples
Total number of samples: __`r nrow(annot_dat)`__   
Number of samples with a known time to recurrence: __`r sum(!is.na(annot_dat$daystorecurrence))`__  
Number of censored samples without a known time to recurrence: __`r sum(is.na(annot_dat$daystorecurrence))`__

```{r, echo = FALSE}
overviewplot <- annot_dat %>%
                arrange(daystoevent) 
overviewplot$censored <- ifelse(overviewplot$censored ==1, FALSE, TRUE)

ggplot(data = overviewplot, aes(x = reorder(Sample,daystoevent), y = daystoevent, alpha = censored, shape = censored, color = Radiotherapy)) +
          geom_point() +
          theme_minimal() +
          theme(axis.text.x = element_text(size = 5, angle = 90)) +
          scale_alpha_manual(values = c(1, 0.4),
                             breaks = c(FALSE, TRUE)) +
          scale_color_manual(values = c("grey50", "red"),
                             breaks = c("No", "Yes")) +
          labs(x = "Sample", y = "Days to Event") 
```

# Cox proportional hazards model

### Overview covariates

Cox proportional hazards model for all clinical parameters that might have an influence on / or are influenced by recurrence-free-survival time.

```{r, echo=FALSE, fig.height= 15}
surv_obj_cox <- Surv(time = annot_dat$daystoevent, event = annot_dat$censored)

fit.coxph_cox <- coxph(surv_obj_cox ~ Radiotherapy + `Tumor Stage` + cluster, data = annot_dat)

plot_overview <- survminer::ggforest(fit.coxph_cox, data = annot_dat)
print(plot_overview)

tiff("Cox_compare_effect.tiff", units = "in", width = 10, height = 5, res = 300, pointsize =(10*200/72))
print(plot_overview)
dev.off()
```

```{r, echo=FALSE, warning=FALSE, fig.height = 8}
surv_obj <- Surv(time = annot_dat$daystoevent, event = annot_dat$censored)
fit2 <- survfit(surv_obj ~ Radiotherapy, data = annot_dat)

ggsurvplot(fit2, data = annot_dat, pval = TRUE, title = "Time-to-Recurrence Radiotherapy", 
                   xlab = "Time-to-Recurrence [days]", ylab = "Recurrence-free survival",
                   palette = c("grey", "red"), risk.table = TRUE)
```

### Protein hits

For each protein/each row in the expression matrix the following steps are performed:

1) Expression data of respective protein is transferred into long format and NAs are removed. 
1b) For non-imputed data, NAs have to be removed more carefully. A maximum fraction of missing values per protein is defined and only proteins above this threshold undergo the statistical test. 

2) A matrix containing the time to event (either recurrence or last follow-up date) for each sample is created. Censoring information is included as well. 
The subsequent Cox regression test checks, whether the hazard of having a recurrence changes depending on the protein's expression level. A hazard ratio and p-values from three different confirmation tests (Likelihood-ratio, Wald, Logrank) are computed. If the hazard ratio is e.g. 0.5, an increase of the selected protein's abundance by 1 halves the risk to suffer from a recurrence at a given time-point. CAVE: in this case, protein abundance means log2(LFQ intensity)!

3) To correct for multiple testing, an adjusted p-value is calculated via Benjamini-Hochberg (FDR based correction). All results are saved as a CSV. 

4) All significant protein hits are summarized in a forest plot. Additionally, a plot is saved for each individual protein. 


```{r, include = FALSE}
adjpvalue_threshold <- 0.05
```

__adjusted p-value threshold = `r adjpvalue_threshold`__  

```{r, include = FALSE}
nathreshold <- 0.2
```

__Maximum fraction of missing values per protein = `r nathreshold`__

```{r, echo=FALSE, include=FALSE}
stats_table <- data.frame(matrix(nrow = 0, ncol = 5))
colnames(stats_table) <- c("ID", "HazardRatio","pvalue", "lower95", "upper95")

i <-  nrow(wide_dat)
while (i > 0) {
  
      protein <- wide_dat[i,] 
      #1)
      obj <-  filter(wide_dat[i,]) %>%
              column_to_rownames(var = "ID") %>%
              pivot_longer(cols = colnames(wide_dat[,-1]), names_to = "Sample", values_to = "Abundance") %>%
              na.omit() %>%
              left_join(annot_dat, by = "Sample")
      
      #1b)
      if (sum(!is.na(obj$Abundance)) >= (1-nathreshold)*(ncol(wide_dat)-1)) {
        
              #2)
              surv_obj <- Surv(time = obj$daystoevent, event = obj$censored)
              fit.coxph <- coxph(surv_obj ~ Abundance + Radiotherapy, data = obj) 
              
              coxph_result <- summary(fit.coxph)
              coxph_pvalue <- coxph_result$coefficients
              coxph_pvalue <- coxph_pvalue[, 5]
              coxph_hazard <- coxph_result$conf.int
              
              
              stats_table <- rbind(stats_table, t(rbind(ID = protein$ID, HazardRatio = coxph_hazard[1,1], 
                                                        pvalue = coxph_pvalue, lower95 = coxph_hazard[1,3], 
                                                        upper95 = coxph_hazard[1,4])))
              }
     
      i <- i-1
      print(i)
}
stats_table$HazardRatio <- as.numeric(stats_table$HazardRatio)
stats_table$pvalue <- as.numeric(stats_table$pvalue)
stats_table$lower95 <- as.numeric(stats_table$lower95)
stats_table$upper95 <- as.numeric(stats_table$upper95)
stats_table <- mutate(stats_table, type = rownames(stats_table)) %>%  #activate if Covariables are included
              filter(str_detect(type, "Abundance") == TRUE)
```

```{r, echo=FALSE, include=FALSE}
#3
stats_table <- mutate(stats_table, adj.pvalue = p.adjust(stats_table$pvalue, method = "BH"))

write.csv(stats_table, "results/stats_table.csv")

adj.hits_stats <- filter(stats_table, adj.pvalue <= adjpvalue_threshold) %>%
                arrange(HazardRatio)
adj.hits <- filter(wide_dat, ID %in% adj.hits_stats$ID)  
```

```{r, echo=FALSE, include=FALSE}
humanproteome <- read.delim("Data/uniprot-proteome UP000005640+reviewed yes(incl.GO).tab") %>%
dplyr::rename(ID = "Entry") 
entries <- str_split_fixed(humanproteome$Entry.name, "_", n=Inf)
genes <- str_split_fixed(humanproteome$Gene.names, " ", n=Inf)
humanproteome$Entry.name <- entries[,1]
humanproteome$Gene.names <- genes[,1]

#remove redundant Uniprot IDs
adj.hits_stats2 <- adj.hits_stats %>%
                   mutate(IDunique = ID)
adj.hits_stats2$ID <- substr(adj.hits_stats$ID, nchar(adj.hits_stats$ID) - 6 + 1, nchar(adj.hits_stats$ID))

adj.hits_stats2 <- left_join(adj.hits_stats2, humanproteome, by = "ID")
                    
write.csv(adj.hits_stats2, file = "results/Coxph_adj.hits_stats.csv")
```

```{r, echo = FALSE, fig.width = 12, fig.height = 8}
ggtable <- adj.hits_stats2 %>%
            mutate(log2HazardRatio = log2(round(HazardRatio, digits = 3))) %>%
            mutate(log2lower95 = log2(lower95)) %>%
            mutate(log2upper95 = log2(upper95)) %>%
  mutate(IDunique = paste(Gene.names)) 
            #mutate(IDunique = paste(IDunique, " / ", Gene.names)) 

#ggtable <- ggtable[-26,] #only if duplicate rows appear (after inclusion of covariables)
ggtable$IDunique <- factor(ggtable$IDunique, levels = ggtable$IDunique)

ggplot(data = ggtable, aes(x = IDunique, y = log2HazardRatio, 
                           ymin = log2lower95, ymax = log2upper95,
                           color = adj.pvalue)) +
  geom_point(size = 3) +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 9, angle = 90, face = "bold")) +
  geom_errorbar() +
  scale_color_gradient(limits = c(0, 0.05), low = "red", high = "purple") +
  labs(x = "Protein Hits (Uniprot / Gene IDs)", y = "log2 Hazard Ratio", color = "p.adjust") 
```

CPHM and Kaplan-Meier Curve for EIF4A1
```{r}
obj <-  filter(wide_dat, ID == "P60842") %>%
              column_to_rownames(var = "ID") %>%
              pivot_longer(cols = 1:ncol(.), names_to = "Sample", values_to = "Abundance") %>%
              na.omit() %>%
              left_join(annot_dat, by = "Sample") %>%
              mutate(EIF4A1 = ifelse(Abundance >= median(Abundance), "high", "low"))

surv_obj <- Surv(time = obj$daystoevent, event = obj$censored)
fit.coxph <- coxph(surv_obj ~ Abundance, data = obj)
              
plot_overview <- ggforest(fit.coxph, data = obj)
print(plot_overview)

tiff("Cox_EIF4A1.tiff", units = "in", width = 6, height = 3, res = 300, pointsize =(6*200/72))
print(plot_overview)
dev.off()

fit16 <- survfit(surv_obj ~ EIF4A1, data = obj)

kapEIF4A1 <- ggsurvplot(fit16, data = obj, pval = TRUE, title = "Time-to-Recurrence EIF4A1", 
                   xlab = "Time-to-Recurrence [days]", ylab = "Recurrence-free survival", 
                  conf.int = FALSE, risk.table = FALSE, palette = c("red", "darkblue"),
            n.line = "v", pval.coord = c(1200, 0.8),
                   pval.method = FALSE)
kapEIF4A1

tiff("KM_EIF4A1.tiff", units = "in", width = 4, height = 3, res = 300, pointsize =(4*200/72))
kapEIF4A1
dev.off()
```


# Gene Ontology of Protein Hits

Different enrichment analyses of protein hits shown above. Performed with GeneOntology and KEGG databases.

### Proteins associated with higher risk

##### Gene Set Enrichment Analysis 

```{r, figure.widh = 12, figure.height = 8, echo = FALSE, message = FALSE, warning = FALSE}
hazardgse <- rename(stats_table, UNIPROT = "ID")
hazardgse$HazardRatio <-  log2(hazardgse$HazardRatio)

gene.df <- bitr(hazardgse$UNIPROT, fromType = "UNIPROT", 
                toType = c("ENTREZID"),
                OrgDb = org.Hs.eg.db)
gseList <- left_join(gene.df, hazardgse, by = "UNIPROT") %>%
              select(ENTREZID, HazardRatio)

geneList = gseList[,2]
names(geneList) = as.character(gseList[,1])
geneList = sort(geneList, decreasing = TRUE)

gse <- gseGO(geneList, ont = "BP", OrgDb = org.Hs.eg.db, keyType = "ENTREZID", minGSSize = 30, maxGSSize = 500, 
               pvalueCutoff = 0.05, pAdjustMethod = "BH", verbose = FALSE)
ridgeplot(gse, label_format = 20) + 
  theme(axis.text.y = element_text(size = 6)) + 
  scale_y_discrete(labels = function(x) str_wrap(x, width = 35)) +
  labs(x = "log2 Hazard Ratio", title = "Ontology: Biological Processes")

gse2 <- gseGO(geneList, ont = "MF", OrgDb = org.Hs.eg.db, keyType = "ENTREZID", minGSSize = 30, maxGSSize = 500, 
               pvalueCutoff = 0.05, pAdjustMethod = "BH", verbose = FALSE)
ridgeplot(gse2, label_format = 20) + 
  theme(axis.text.y = element_text(size = 6)) +
  labs(x = "log2 Hazard Ratio", title = "Ontology: Molecular Function")
```


GSEA of all proteins associated with higher risk to recurrence in the dataset: Hazard Ratio > 1 (no p-value adjustment)

```{r, figure.widh = 12, echo = FALSE, message = FALSE, warning = FALSE}
hazardupgse <- filter(stats_table, stats_table$HazardRatio > 1) %>%
              rename(UNIPROT = "ID")
hazardupgse$HazardRatio <-  log2(hazardupgse$HazardRatio)

gene.df_up <- bitr(hazardupgse$UNIPROT, fromType = "UNIPROT", 
                toType = c("ENTREZID"),
                OrgDb = org.Hs.eg.db)
gseListup <- left_join(gene.df_up, hazardupgse, by = "UNIPROT") %>%
              select(ENTREZID, HazardRatio)

geneListup = gseListup[,2]
names(geneListup) = as.character(gseListup[,1])
geneListup = sort(geneListup, decreasing = TRUE)

gse_up <- gseGO(geneListup, ont = "BP", OrgDb = org.Hs.eg.db, keyType = "ENTREZID", minGSSize = 30, maxGSSize = 500, 
               pvalueCutoff = 0.05, pAdjustMethod = "BH", verbose = FALSE)
dotplot(gse_up, label_format =20, font.size = 6) + 
  theme(axis.text.y = element_text(size = 8)) +
  labs(title = "Ontology: Biological Process", subtitle = "Hazard Ratio > 1")

gse_up2 <- gseGO(geneListup, ont = "MF", OrgDb = org.Hs.eg.db, keyType = "ENTREZID", minGSSize = 30, maxGSSize = 500, 
               pvalueCutoff = 0.05, pAdjustMethod = "BH", verbose = FALSE)
cnetplot(gse_up2, node_label="category", categorySize = "geneNum", foldChange = gse_up2, cex_label_category = 1)
```
  
### Proteins associated with lower risk

##### Gene Set Enrichment Analysis 

GSEA of all proteins associated with lower risk to recurrence in the dataset: Hazard Ratio < 1 (no p-value adjustment)

```{r, figure.widh = 12, echo = FALSE, message = FALSE, warning = FALSE}
hazarddowngse <-  filter(stats_table, stats_table$HazardRatio < 1) %>%
                  rename(UNIPROT = "ID") 
hazarddowngse <- rename(stats_table, UNIPROT = "ID")
hazarddowngse$HazardRatio <-  log2(hazarddowngse$HazardRatio)  

gene.df_down <- bitr(hazarddowngse$UNIPROT, fromType = "UNIPROT", 
                toType = c("ENTREZID"),
                OrgDb = org.Hs.eg.db)
gseListdown <- left_join(gene.df_down, hazarddowngse, by = "UNIPROT") %>%
              select(ENTREZID, HazardRatio)

geneListdown = gseListdown[,2]
names(geneListdown) = as.character(gseListdown[,1])
geneListdown = sort(geneListdown, decreasing = TRUE)

gse_down <- gseGO(geneListdown, ont = "MF", OrgDb = org.Hs.eg.db, keyType = "ENTREZID", minGSSize = 30, maxGSSize = 500, 
               pvalueCutoff = 0.05, pAdjustMethod = "BH", verbose = FALSE)

dotplot(gse_down, label_format =20, font.size = 6) + 
  theme(axis.text.y = element_text(size = 8)) +
  labs(title = "Ontology: Biological Process", subtitle = "Hazard Ratio < 1")

cnetplot(gse_down, node_label="category", categorySize = "geneNum", foldChange = gse_up2, cex_label_category = 1)
```


